---
title: "Code-book"
output: html_document
---


Code Book

This code book summarizes the resulting data fields in tidy.txt.
Identifiers

    subject - id of subject
    activity_code - Activity type 


Activity Labels

    1 - WALKING: subject was WALKING during the test
    2 - WALKING_UPSTAIRS: subject was WALKING_UPSTAIRS during the test
    3 - WALKING_DOWNSTAIRS: subject was WALKING_DOWNSTAIRS during the test
    4 - SITTING: subject was SITTING during the test
    5 - STANDING: subject was STANDING during the test
    6 - LAYING: subject was LAYING down during the test
    
Variables (Measurements)

    tBodyAccMeanX
    tBodyAccMeanY
    tBodyAccMeanZ
    tBodyAccStdX
    tBodyAccStdY
    tBodyAccStdZ
    tGravityAccMeanX
    tGravityAccMeanY
    tGravityAccMeanZ
    tGravityAccStdX
    tGravityAccStdY
    tGravityAccStdZ
    tBodyAccJerkMeanX
    tBodyAccJerkMeanY
    tBodyAccJerkMeanZ
    tBodyAccJerkStdX
    tBodyAccJerkStdY
    tBodyAccJerkStdZ
    tBodyGyroMeanX
    tBodyGyroMeanY
    tBodyGyroMeanZ
    tBodyGyroStdX
    tBodyGyroStdY
    tBodyGyroStdZ
    tBodyGyroJerkMeanX
    tBodyGyroJerkMeanY
    tBodyGyroJerkMeanZ
    tBodyGyroJerkStdX
    tBodyGyroJerkStdY
    tBodyGyroJerkStdZ
    tBodyAccMagMean
    tBodyAccMagStd
    tGravityAccMagMean
    tGravityAccMagStd
    tBodyAccJerkMagMean
    tBodyAccJerkMagStd
    tBodyGyroMagMean
    tBodyGyroMagStd
    tBodyGyroJerkMagMean
    tBodyGyroJerkMagStd
    fBodyAccMeanX
    fBodyAccMeanY
    fBodyAccMeanZ
    fBodyAccStdX
    fBodyAccStdY
    fBodyAccStdZ
    fBodyAccMeanFreqX
    fBodyAccMeanFreqY
    fBodyAccMeanFreqZ
    fBodyAccJerkMeanX
    fBodyAccJerkMeanY
    fBodyAccJerkMeanZ
    fBodyAccJerkStdX
    fBodyAccJerkStdY
    fBodyAccJerkStdZ
    fBodyAccJerkMeanFreqX
    fBodyAccJerkMeanFreqY
    fBodyAccJerkMeanFreqZ
    fBodyGyroMeanX
    fBodyGyroMeanY
    fBodyGyroMeanZ
    fBodyGyroStdX
    fBodyGyroStdY
    fBodyGyroStdZ
    fBodyGyroMeanFreqX
    fBodyGyroMeanFreqY
    fBodyGyroMeanFreqZ
    fBodyAccMagMean
    fBodyAccMagStd
    fBodyAccMagMeanFreq
    fBodyBodyAccJerkMagMean
    fBodyBodyAccJerkMagStd
    fBodyBodyAccJerkMagMeanFreq
    fBodyBodyGyroMagMean
    fBodyBodyGyroMagStd
    fBodyBodyGyroMagMeanFreq
    fBodyBodyGyroJerkMagMean
    fBodyBodyGyroJerkMagStd
    fBodyBodyGyroJerkMagMeanFreq


Work:

The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set.
Review criteria

    - The submitted data set is tidy.
    - The Github repo contains the required scripts.
    - GitHub contains a code book that modifies and updates the available codebooks with the data to indicate all the variables and
      summaries calculated, along with units, and any other relevant information.
    - The README that explains the analysis files is clear and understandable.
    - The work submitted for this project is the work of the student who submitted it.

Getting and Cleaning Data Course Project

The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 

    - 1) a tidy data set as described below, 
    - 2) a link to a Github repository with your script for performing the analysis, and 
    - 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called
         CodeBook.md. 
         
         You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are 
         connected.

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

Here are the data for the project:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

You should create one R script called run_analysis.R that does the following.

    - Merges the training and the test sets to create one data set.
    - Extracts only the measurements on the mean and standard deviation for each measurement.
    - Uses descriptive activity names to name the activities in the data set
    - Appropriately labels the data set with descriptive variable names.
    - From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each
      subject.

Good luck!

```{r}
library(dplyr)

# Downloading & Unzipping Folder
url = "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
filename = "dataset.zip"
download.file(url,filename,method="curl")
unzip(filename)
```

```{r}
# Reading Activity Labels & Features 
activity_labels <- read.table("UCI HAR Dataset/activity_labels.txt")
features <- read.table("UCI HAR Dataset/features.txt")

# Reading training set
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt")
colnames(subject_train) = features$V2

x_train <- read.table("UCI HAR Dataset/train/X_train.txt")
colnames(x_train) = features$V2

y_train <- read.table("UCI HAR Dataset/train/y_train.txt")
colnames(y_train) = c("activity_code")

# Reading test set
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt")
colnames(subject_test) = c("subject")

x_test <- read.table("UCI HAR Dataset/test/X_test.txt")
colnames(x_test) = features$V2

y_test <- read.table("UCI HAR Dataset/test/y_test.txt")
colnames(y_test) = c("activity_code")
```


1. Merges the training and the test sets to create one data set.

```{r}
x = rbind(x_train,x_test)
y = rbind(y_train,y_test)
subject = rbind(subject_train,subject_test)

df = cbind(subject,y,x)
head(df)
```

2. Extracts only the measurements on the mean and standard deviation for each measurement.
```{r}
df2 <- df[,grep("std\\(\\)|mean\\(\\)|activity_code|subject", colnames(df), value=TRUE)] 
head(df2)
```

- Uses descriptive activity names to name the activities in the data set
```{r}
df2$activity_description = factor(df2$activity_code, levels = activity_labels[,1], labels = activity_labels[,2])
head(df2)
```

- Appropriately labels the data set with descriptive variable names.
```{r}
names(df2)<-gsub("Acc", "Accelerometer", names(df2))
names(df2)<-gsub("Gyro", "Gyroscope", names(df2))
names(df2)<-gsub("BodyBody", "Body", names(df2))
names(df2)<-gsub("Mag", "Magnitude", names(df2))
names(df2)<-gsub("^t", "Time", names(df2))
names(df2)<-gsub("^f", "Frequency", names(df2))
names(df2)<-gsub("tBody", "timeBody", names(df2))
names(df2)<-gsub("-mean()", "Mean", names(df2), ignore.case = TRUE)
names(df2)<-gsub("-std()", "Std", names(df2), ignore.case = TRUE)
names(df2)<-gsub("-freq()", "Freq", names(df2), ignore.case = TRUE)
names(df2)<-gsub("[-()]", "", names(df2))
head(df2)
```

- From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each
  subject.
  
```{r}
final_df <- df2 %>%
  group_by(subject, activity_code) %>%
  summarise_all(funs(mean))
  
head(final_df)
write.table(final_df, "final_df.txt", row.name=FALSE)
```

  


